A survey of HTTP caching options on the open Semantic Web

### 2015-01-19 sanity/paper1.tex
Improve methodology; promising voes
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@1013 43ef1190-90fd-4dd7-a769-d0121f02b80e

RFC7232, on the other hand, defines a protocol for asking the server
if the cached response is still fresh using conditional requests. This
doesn't burden the content provider with the task of estimating the
time to live beforehand, but then, it must be able to answer if the
resource has changed less expensively than serving the entire
response. Either of two headers must be set by the server to achieve
this: \httph{ETag}, which sets an opaque identifier for the response,
or \httph{Last-Modified} which gives the time and date of the last
modification of the resource. Clients may use these values to retrieve
the updated response only if the resource has changed. Other relevant
headers are described in Table~\ref{tab:headers}
in Appendix~\ref{app:fetcher} but they were not all used in this
analysis.

@@

\section{Methodology}
\todo{Improve, big}
We want to find information resources on the Web, and examine HTTP
headers that may allow caching. We record headers recommended by
current standards, as well as obsoleted and non-standard
headers. Additionally, we examine the triples in the returned
information resources to see if there is information there that may be
used to calculate heuristic freshness.

However, as the community has both spidered the Web as well as
recorded lists of relevant resources, we need not do that ourselves.

The SPARQLES survey\cite{buil2013sparql} has recorded the availability
of SPARQL endpoints registered in the datahub.io portal, and found
that many of them have ceased to operate. We used their list of SPARQL
endpoints as of 2014-11-17, and filtered out those deemed unresponsive
by them. This resulted in a list of 312 endpoints.

To compile a list of Web ontologies, we first queried the Linked Open
Vocabularies (LOV) \cite{lov2} SPARQL endpoint
to retrieve all registered ontologies. Then, we took the full list of
URIs registered with prefix.cc. The assumption that they all identify
an ontology is clearly false, as we found that for example Tim
Berners-Lee's FOAF profile has a registered URI, but it is unlikely to
introduce significant bias.

To examine as many different implementations and hosts as possible, we
noted that the Billion Triple Challenge 2014 \cite{btc-2014} dataset
consisted of a 4~GTriple corpus of spidered Web data. To compile a
list of interesting candidates to further examine, we performed a
series of data reduction steps, manually inspecting the result between
each step. The details of this process is given in
Appendix~\ref{app:reduction}.

The end result of this process is a list of 3117 unique hosts, for
each several resources would be visited, some several times, as they
may host SPARQL Endpoints, ontologies, or other information resources,
by a spider detailed in Appendix~\ref{app:fetcher}, resulting in 7745
requests, done on 2015-01-02.

This results in an NQuads file per host, which is then loaded into a
Virtuoso-based SPARQL Endpoint for analysis by using the statistics
system R\cite{kn:r}.

@@

First, we investigated whether certain server implementations provided better support
for caching than others. To do this, we formulated SPARQL queries to
examine the \httph{Server} headers of successful responses, with
optional clauses matching the standard-compliant computed freshness
time (which is the ideal) as well as whether the response had other
promising features that could assist caching, modification time,
certain predicates, etc.

For each unique \httph{Server} header, we found the ones where
\emph{all} responses had a freshness time or a promising\todo{what are
  they} feature. For the former, this amounted to 22 servers. 70
servers always responded with promising features. Inspecting the
values we find the well-known Virtuoso and Callimachus servers, as
well as the Perl modules RDF::LinkedData and RDF::Endpoint, which is
partly developed by and runs on a server operated by this
author. Apart from those, a typical server string is
e.g. \httph{Apache/2.2.22 (Fedora)}, which reveals very little about
the RDF-specific parts of the underlying server implementation. A
quick inspection of all \httph{Server} headers confirmed that few
reveal any further detail.

@@

Of these resources, 405 returned valid headers, but 114 did so to
prohibit caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching.

Common practice when prohibiting caching is to set the \httph{Expires}
header to a date far back in time, which has also been done in large
parts of the present dataset. For the following, we have set any
non-positive lifetime to 0, as that is semantically equivalent.

@@

We then sought to classify resources into the categories ``endpoint''
for SPARQL endpoints, ``vocabulary'' for ontologies, ``dataset'' for
datasets that may contain further descriptions of several resources,
or simply ``inforesources'' for those that did not fit in the above
classes. To do so, we classified based on certain predicates and
classes. Additionally, URIs derived from prefix.cc were classified as
``vocabulary'' and those from SPARQLES as ``endpoint''.

>>>>

RFC7232, on the other hand, defines a protocol for asking the server
if the cached response is still fresh using conditional requests. This
doesn't burden the content provider with the task of estimating the
time to live beforehand, but then, it must be able to answer if the
resource has changed less expensively than serving the entire
response. Either of two headers must be set by the server to achieve
this: \httph{ETag}, which sets an opaque identifier for the response,
or \httph{Last-Modified} which gives the time and date of the last
modification of the resource. Clients may use these values to retrieve
the updated response only if the resource has changed. Other relevant
headers are described in Table~\ref{tab:headers}
in Appendix~\ref{app:fetcher}.

@@

\section{Methodology}
We want to find information resources on the Web, and examine HTTP
headers that may allow caching. To do this, we perform GET requests on
SPARQL endpoints, vocabularies, dataset descriptions and other
resources and record headers recommended by current standards, as well
as obsoleted and non-standard headers. Additionally, we examine the
triples in the returned information resources to see if there is
information there that may be used to calculate heuristic freshness.

We make several approaches to ensure that we visit a large and
representative section of the open Semantic Web. We take SPARQL
Endpoints from the SPARQLES survey \cite{buil2013sparql}, vocabularies
from Linked Open Vocabularies (LOV) \cite{lov2} and prefix.cc, and
we augment these data with spidered data from the Billion Triple
Challenge 2014 \cite{btc-2014} dataset.

We used SPARQLES survey list of SPARQL endpoints as of 2014-11-17, and
filtered out those deemed unresponsive by them. This resulted in a
list of 312 endpoints.

To examine as many different implementations and hosts as possible, we
noted that the Billion Triple Challenge 2014 \cite{btc-2014} dataset
consisted of a 4~GTriple corpus of spidered Web data. To compile a
list of candidates to further examine, we performed a
series of data reduction steps, manually inspecting the result between
each step. The details of this process is given in
Appendix~\ref{app:reduction}.

The end result of this process is a list of 3117 unique hosts, for
each several resources would be visited, some several times, as they
may host SPARQL Endpoints, vocabularies, or other information
resources, by a spider detailed in Appendix~\ref{app:fetcher},
resulting in 7745 requests, done on 2015-01-02. Although this is not a
very large number, we assume that caching headers are usually set on a
per-host basis, and so the number of hosts is more important than the
number of resources.

This results in an NQuads file per host, which is then loaded into a
Virtuoso-based SPARQL Endpoint for analysis by using the statistics
system R\cite{kn:r} in the following section.

@@

First, we investigated whether certain server implementations provided better support
for caching than others. To do this, we formulated SPARQL queries to
examine the \httph{Server} headers of successful responses, with
optional clauses matching the standard-compliant computed freshness
lifetime (which is the ideal) as well as whether the response had other
features that may assist caching such as modification time,
certain predicates, etc.\todo{promising}

For each unique \httph{Server} header, we found the ones where
\emph{all} responses had a freshness time or a promising feature. For
the former, this amounted to 22 servers. 70 servers always responded
with promising features. Inspecting the values we find the well-known
Virtuoso and Callimachus servers, as well as the Perl modules
RDF::LinkedData and RDF::Endpoint, which is partly developed by and
runs on a server operated by this author. Apart from those, a typical
server string is e.g. \httph{Apache/2.2.22 (Fedora)}, which reveals
very little about the RDF-specific parts of the underlying server
implementation. A quick inspection of all \httph{Server} headers
confirmed that few reveal any further detail.

@@

Of the 2965 resources, 405 returned valid headers, but 114 did so to
prohibit caching of the response, and 3 contained conflicting
headers, i.e. set an freshness lifetime, but also prohibited
caching. In most cases, \httph{Cache-Control} and \httph{Expires} both
occurred, but the former is more common than the latter in the cases
where only one of them occur.  Additionally, 269 resources had a
\httph{Cache-Control} header to control other aspects of caching than
lifetime, i.e. to say that only private caches may use the response,
that the cache must be revalidated, or to prohibit caching.

% todo: kill this?
Common practice when prohibiting caching is to set the \httph{Expires}
header to a date far back in time, which has also been done in large
parts of the present dataset. For the following, we have set any
non-positive lifetime to 0, as that is semantically equivalent.

@@

We then sought to classify resources into the categories ``endpoint''
for SPARQL endpoints, ``vocabulary'' for vocabularies, ``dataset'' for
datasets that may contain further descriptions of several resources,
or simply ``inforesources'' for those that did not fit in the above
classes. To do so, we classified based on certain predicates and
classes. Additionally, URIs derived from prefix.cc were classified as
``vocabulary'' (even if we found several that were not) and those from
SPARQLES as ``endpoint''.
